---
title: "Retail Trends"
---

```{r message = FALSE}
library(tidyverse)
library(lubridate)
library(stringr)

retail <- readRDS("../data/retail_clean.rds")
coords <- readRDS("../data/cty_coords.rds")
features <- readRDS("../data/cust_features.rds")
descriptions <- readLines("../data/column_description.txt")
```

### How does revenue change across time?

```{r fig.align = "center"}
monthly_rev <- retail %>%
  mutate(InvoiceDate, ym = strftime(InvoiceDate, "%Y-%m")) %>%
  group_by(ym) %>% 
  summarise(revenue = sum(UnitPrice*Quantity) / 1000, 
            num_days = as.numeric(max(InvoiceDate) - min(InvoiceDate)), .groups = "drop") %>%
  mutate(project = c("yes", rep("no", 11), "yes"), 
         projected = ifelse(project == "yes", revenue / num_days * 31, revenue))

ggplot(monthly_rev) +
  geom_point(mapping = aes(ym, projected, colour = project)) + 
  scale_color_manual(values = c("yes" = "red", "no" = "black")) +
  geom_line(mapping = aes(ym, projected, group = 1)) + 
  labs(title = "Monthly revenue", x= "Year/Month", y = "Revenue", 
       subtitle = "Decembers are projected") +
  theme(legend.position = "blank")
```

### Rumour has it that Christmas shopping starts from around November. Is that true?

```{r fig.align = "center"}
before_nov <- retail %>%
  filter(InvoiceDate < date("2011-11-01") & InvoiceDate > date("2011-01-01")) %>%
  group_by(CustomerID, InvoiceNo) %>%
  summarise(amt = sum(Quantity * UnitPrice), .groups = "drop_last") %>%
  slice_max(InvoiceNo, n = 3) %>% 
  summarise(mean_amt = mean(amt), .groups = "drop")

after_nov <- retail %>%
  filter(InvoiceDate >= date("2011-11-01") & InvoiceDate < date("2012-01-01")) %>%
  group_by(CustomerID, InvoiceNo) %>%
  summarise(amt = sum(Quantity * UnitPrice), .groups = "drop_last") %>%
  slice_min(InvoiceNo, n = 3) %>% 
  summarise(mean_amt = mean(amt), .groups = "drop") 

before_after_nov <- left_join(before_nov, after_nov, by = c("CustomerID" = "CustomerID")) %>%
  filter(!is.na(mean_amt.y)) %>%
  rename(mean_amt_before_nov = mean_amt.x, mean_amt_after_nov = mean_amt.y) %>%
  pivot_longer(cols = c(mean_amt_before_nov, mean_amt_after_nov), names_to = "time", values_to = "mean_amt_spent")

ggplot(before_after_nov)  +
  geom_boxplot(mapping = aes(time, mean_amt_spent))
```

Hmm... doesn't seem to be true. The distributions seem quite even. What if we only looked at Christmas related items?

```{r fig.align = "center"}
christmas_item_track <- mutate(retail, Date = date(InvoiceDate), 
                               christmas_item = ifelse(str_detect(Description, "(CHRISTMAS|XMAS|SANTA|REINDEER)"), 1, 0)) %>%
  group_by(Date) %>%
  summarise(num_items = sum(christmas_item), .groups = "drop")

ggplot(christmas_item_track) + 
  geom_line(mapping = aes(x = Date, y = num_items)) +
  labs(title = "Christmas items trend", y = "Christmas items purchased")

```

Ahh yes. As expected, Christmas items begin selling off the shelves closer to December.

### Extending from the first question, how does each country perform across the quarters? 

```{r fig.align = "center", fig.height = 6, fig.width = 10}
retail <- mutate(retail, year = year(InvoiceDate), month = month(InvoiceDate))

country_per_quarter <- retail %>%
  filter(year == 2011) %>%
  mutate(quarter = case_when(
    between(month, 1, 3) ~ 1,
    between(month, 4, 6) ~ 2,
    between(month, 7, 9) ~ 3,
    TRUE ~ 4
  )) %>%
  group_by(Country, quarter) %>%
  summarise(mean_revenue = mean(Quantity * UnitPrice), .groups = "drop") %>%
  left_join(coords, by = c("Country" = "rtl"))

  to_reorder <- country_per_quarter %>%
  group_by(Country) %>%
  summarise(total = sum(mean_revenue), .groups = "drop") %>%
  mutate(Country = reorder(Country, total))

country_per_quarter <- mutate(country_per_quarter, Country = factor(Country, levels = rev(levels(to_reorder$Country)))) %>%
  filter(Country != "Unspecified")

ggplot(country_per_quarter) +
  geom_col(mapping = aes(quarter, mean_revenue, fill = group)) + 
  facet_wrap(~ Country) +
  labs(title = "How does each country perform across the quarters?", x = "Quarter", y = "Mean revenue") + 
  scale_fill_discrete(name = "Region")
  
```

Seems like the Asian countries dominate.

### How does the median unique items purchased by customers from each country vary from one place to another? 

```{r fig.align = "center", fig.height = 6, fig.width = 8}
med_unique_items <- filter(features, grp2 != "NA") %>%
  group_by(grp2, Country) %>%
  summarise(med_unique = median(m_uq), .groups = "drop") %>%
  mutate(Country = reorder(Country, med_unique))
ggplot(med_unique_items) +
  geom_point(mapping = aes(Country, med_unique, colour = grp2), size = 4) + 
  coord_flip() +
  labs(title = "Median unique purchases per invoice in countries", 
       y = "Median unique purchases per invoice", fill = "Region")
```

There seems to be a pattern...some countries that rank higher on the graph for `country_per_quarter` seem to rank lower on the graph for `med_unique_items`. Let's see if this is actually the case.

```{r}
med_u_rank <- mutate(med_unique_items, med_u_items_rank = min_rank(desc(med_unique)))
mean_rev_rank <- mutate(to_reorder, sum_mean_rev_rank = min_rank(desc(total)))
correl <- left_join(med_u_rank, mean_rev_rank) %>%
  select(med_u_items_rank, sum_mean_rev_rank) %>%
  cor(use = "pair") %>%
  as.dist() 
correl
```

A negative correlation! That suggests that higher expenditure can be attributed to buying more of the same items. 
